MSstatsPTM_SImulation_Analysis

### 2022-01-28 code/MSstatsPTM_Simulation_Analysis.Rmd
Changes for internal review

## Label Models
adj_results$Model <- "Proposed"
no_adj_results$Model <- "Proposed no Adjustment"
ttest_results$Model <- "T-Test no Adjustment"
ttest_adj_results$Model <- "T-Test"
limma_model_results$Model <- "Limma no Adjustment"
limma_adj_model_results$Model <- "Limma"

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width = 7, fig.height = 6}
png("../supplementary/sim_new/sim1_FDR_all_models.png", width = 750, height = 500)

models %>% mutate(Model = factor(Model, levels=c("Proposed", "Proposed no Adjustment", "Limma", 
                                               "Limma no Adjustment", "T-Test", "T-Test no Adjustment")),
                  Replicates = as.factor(rep)) %>%
  ggplot() + geom_point(aes(x = Model, y = fpr, size = Replicates, color = Model)) + 
  scale_colour_manual(values=cbPalette_first) +
  facet_grid(vars(sd), vars(conditions)) + 
  theme_bw() +
  theme(axis.text.x = element_blank(), 
                                                 axis.text.y = element_text(size = 12), 
                                                 legend.text=element_text(size=12),
                                                 axis.title.y = element_text(size = 14),
                                                 axis.title.x = element_text(size = 14),
                                                 title = element_text(size = 16),
                                                 strip.text = element_text(size = 12)) + 
  labs(title = "Simulation 1: FDR all models", x = "Model", y = "FDR")

@@

``` {r, echo = FALSE, message=FALSE, warning=FALSE}
png("../supplementary/sim_new/sim1_FDR.png", width = 750, height = 500)

models %>% mutate(Model = factor(Model, levels=c("Proposed", "Proposed no Adjustment", "Limma", 
                                               "Limma no Adjustment", "T-Test", "T-Test no Adjustment")),
                  Replicates = as.factor(rep)) %>%
  filter(Model %in% c("Proposed", "T-Test", "Limma")) %>% 
  ggplot() + geom_point(aes(x = Model, y = fpr, size = Replicates, color = Model)) + 
  scale_colour_manual(values=cbPalette) +
  facet_grid(vars(sd), vars(conditions)) + 
  theme_bw() +
  theme(axis.text.x = element_blank(), 
                                                 axis.text.y = element_text(size = 12), 
                                                 legend.text=element_text(size=12),
                                                 axis.title.y = element_text(size = 14),
                                                 axis.title.x = element_text(size = 14),
                                                 title = element_text(size = 16),
                                                 strip.text = element_text(size = 12)) + 
  labs(title = "Simulation 1: FDR adjusted models", x = "Model", y = "FDR")
dev.off()
```

@@

```{r echo=FALSE, message=FALSE, warning=FALSE}
png("../supplementary/sim_new/sim1_Recall.png", width = 750, height = 500)
models %>% mutate(Model = factor(Model, levels=c("Proposed", "Proposed no Adjustment", "Limma", 
                                               "Limma no Adjustment", "T-Test", "T-Test no Adjustment")),
                  Replicates = as.factor(rep)) %>%
  filter(Model %in% c("Proposed", "T-Test", "Limma")) %>% 
  ggplot() + geom_point(aes(x = Model, y = recall, size = Replicates, color = Model)) + 
  scale_colour_manual(values=cbPalette) +
  facet_grid(vars(sd), vars(conditions)) + 
  theme_bw() +
  theme(axis.text.x = element_blank(), 
                                                 axis.text.y = element_text(size = 12), 
                                                 legend.text=element_text(size=12),
                                                 axis.title.y = element_text(size = 14),
                                                 axis.title.x = element_text(size = 14),
                                                 title = element_text(size = 16),
                                                 strip.text = element_text(size = 12)) + 
  labs(title = "Simulation 1: Recall adjusted models", x = "Model", y = "Recall")
dev.off()
```

@@

```{r echo=FALSE, message=FALSE, warning=FALSE}
png("../supplementary/sim_new/sim1_accuracy.png", width = 750, height = 500)
models %>% mutate(Model = factor(Model, levels=c("Proposed", "Proposed no Adjustment", "Limma", 
                                               "Limma no Adjustment", "T-Test", "T-Test no Adjustment")),
                  Replicates = as.factor(rep)) %>%
  filter(Model %in% c("Proposed", "T-Test", "Limma")) %>% 
  ggplot() + geom_point(aes(x = Model, y = accuracy, size = Replicates, color = Model)) + 
  scale_colour_manual(values=cbPalette) +
  facet_grid(vars(sd), vars(conditions)) + 
  theme_bw() +
  theme(axis.text.x = element_blank(), 
                                                 axis.text.y = element_text(size = 12), 
                                                 legend.text=element_text(size=12),
                                                 axis.title.y = element_text(size = 14),
                                                 axis.title.x = element_text(size = 14),
                                                 title = element_text(size = 16),
                                                 strip.text = element_text(size = 12)) + 
  labs(title = "Simulation 1: Accuracy adjusted models", x = "Model", y = "Accuracy")

@@

png("../supplementary/sim_new/sim1_FC_boxplot.png", width = 750, height = 500)
boxplot_data %>% mutate(Model = factor(Model, levels=c("Proposed", "Limma", "T-test")),
                        reps = factor(reps, levels = c(2,3,5,10))) %>% 
  ggplot()  + geom_hline(yintercept = .75, color = "red", size = 1.1) + 
  geom_boxplot(aes(y = log2FC, x = reps, fill = Model)) + facet_grid(vars(sd), vars(cond)) + 
  scale_fill_manual(values=cbPalette) +
  theme_bw() +
  theme(axis.text.x = element_text(size = 12), 
                                                 axis.text.y = element_text(size = 12), 
                                                 legend.text=element_text(size=12),
                                                 axis.title.y = element_text(size = 14),
                                                 axis.title.x = element_text(size = 14),
                                                 title = element_text(size = 16),
                                                 strip.text = element_text(size = 12)) + 
  labs(title = "Simulation 1: Fold Change Distribution", x = "Replicates", y = "Log2FC")
dev.off()
```

@@

load(file = "model_data/ptm_models_sim2.rda")#ptm_models_sim1
load(file = "model_data/adjusted_models_sim2.rda")#adjusted_models_sim1 
load(file = "model_data/ttest_models_sim2.rda")#ttest_sim1
load(file = "model_data/adj_ttest_models_sim2.rda")#adj_ttest_sim1
load(file = "model_data/limma_models_sim2.rda")#limma_results_sim1
load(file = "model_data/adj_limma_models_sim2.rda")#adj_limma_sim1

@@

``` {r, echo = FALSE, message=FALSE, warning=FALSE}
png("../supplementary/sim_new/sim3_FDR.png", width = 350, height = 350)
models %>% mutate(Model = factor(Model, levels=c("Proposed", "Proposed_no_adj", "limma_no_adj", 
                                               "Limma", "ttest_no_adj", "T-Test")),
                  Replicates = as.factor(rep)) %>%
  filter(Model %in% c("Proposed", "T-Test", "Limma")) %>% 
  ggplot() + geom_point(aes(x = Model, y = fpr, size = Replicates, color = Model)) + 
  scale_colour_manual(values=cbPalette) +
  facet_grid(vars(sd), vars(conditions)) + 
  theme_bw() +
  theme(axis.text.x = element_blank(), 
                                                 axis.text.y = element_text(size = 12), 
                                                 legend.text=element_text(size=12),
                                                 axis.title.y = element_text(size = 14),
                                                 axis.title.x = element_text(size = 14),
                                                 title = element_text(size = 16),
                                                 strip.text = element_text(size = 12)) + 
  labs(title = "FDR adjusted models", x = "Model", y = "FDR")
dev.off()
```

@@

```{r echo=FALSE, message=FALSE, warning=FALSE}
png("../supplementary/sim_new/sim3_Recall.png", width = 750, height = 500)
models %>% mutate(Model = factor(Model, levels=c("Proposed", "Proposed_no_adj", "limma_no_adj", 
                                               "Limma", "ttest_no_adj", "T-Test")),
                  Replicates = as.factor(rep)) %>%
  filter(Model %in% c("Proposed", "T-Test", "Limma")) %>% 
  ggplot() + geom_point(aes(x = Model, y = recall, size = Replicates, color = Model)) + 
  scale_colour_manual(values=cbPalette) +
  facet_grid(vars(sd), vars(conditions)) + 
  theme_bw() +
  theme(axis.text.x = element_blank(), 
                                                 axis.text.y = element_text(size = 12), 
                                                 legend.text=element_text(size=12),
                                                 axis.title.y = element_text(size = 14),
                                                 axis.title.x = element_text(size = 14),
                                                 title = element_text(size = 16),
                                                 strip.text = element_text(size = 12)) + 
  labs(title = "Simulation 2: Recall adjusted models", x = "Model", y = "Recall")
dev.off()
```

@@

```{r echo=FALSE, message=FALSE, warning=FALSE}
png("../supplementary/sim_new/sim3_Accuracy.png", width = 750, height = 500)
models %>% mutate(Model = factor(Model, levels=c("Proposed", "Proposed_no_adj", "limma_no_adj", 
                                               "Limma", "ttest_no_adj", "T-Test")),
                  Replicates = as.factor(rep)) %>%
  filter(Model %in% c("Proposed", "T-Test", "Limma")) %>% 
  ggplot() + geom_point(aes(x = Model, y = accuracy, size = Replicates, color = Model)) + 
  scale_colour_manual(values=cbPalette) +
  facet_grid(vars(sd), vars(conditions)) + 
  theme_bw() +
  theme(axis.text.x = element_blank(), 
                                                 axis.text.y = element_text(size = 12), 
                                                 legend.text=element_text(size=12),
                                                 axis.title.y = element_text(size = 14),
                                                 axis.title.x = element_text(size = 14),
                                                 title = element_text(size = 16),
                                                 strip.text = element_text(size = 12)) + 
  labs(title = "Simulation 2: Accuracy adjusted models", x = "Model", y = "Accuracy")
dev.off()
```

@@

png("../supplementary/sim_new/sim3_FC_boxplot.png", width = 750, height = 500)
boxplot_data %>% mutate(Model = factor(Model, levels=c("Proposed", "Limma", "T-test")),
                        reps = factor(reps, levels = c(2,3,5,10))) %>% 
ggplot()  + geom_hline(yintercept = .75, color = "red", size = 1.1) + 
geom_boxplot(aes(y = log2FC, x = reps, fill = Model)) + facet_grid(vars(sd), vars(cond)) + 
    scale_fill_manual(values=cbPalette) +
  theme_bw() +
  theme(axis.text.x = element_text(size = 12), 
                                               axis.text.y = element_text(size = 12), 
                                               legend.text=element_text(size=12),
                                               axis.title.y = element_text(size = 14),
                                               axis.title.x = element_text(size = 14),
                                               title = element_text(size = 16),
                                               strip.text = element_text(size = 12)) + 
labs(title = "Simulation 2: Fold Change Distribution", x = "Replicates", y = "Log2FC")
dev.off()
```

>>>>

## Label Models
adj_results$Model <- "Proposed"
no_adj_results$Model <- "Proposed no adjustment"
ttest_results$Model <- "T-Test no adjustment"
ttest_adj_results$Model <- "T-Test"
limma_model_results$Model <- "Limma no adjustment"
limma_adj_model_results$Model <- "Limma"

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width = 7, fig.height = 6}

models$plot_conditions = paste0(models$conditions, " Conditions")
models$plot_sd = paste0("SD ", models$sd)

png("../supplementary/sim_new/sim1_FDR_all_models.png", width = 750, height = 500)

models %>% mutate(Model = factor(Model, levels=c("Proposed", "Proposed no adjustment", "Limma", 
                                               "Limma no adjustment", "T-Test", "T-Test no adjustment")),
                  Replicates = as.factor(rep)) %>%
  ggplot() + geom_point(aes(x = Model, y = fpr, size = Replicates, color = Model)) + 
  scale_colour_manual(values=cbPalette_first) +
  scale_size_manual(values = c(3,5,7,9)) + 
  facet_grid(vars(plot_sd), vars(plot_conditions)) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 60, hjust=1, size = 12), 
         axis.text.y = element_text(size = 14), 
         legend.text=element_text(size=14),
         axis.title.y = element_text(size = 22),
         axis.title.x = element_text(size = 22),
         title = element_text(size = 22),
         strip.text = element_text(size = 16)) + 
  labs(title = "Simulation 1: FDR all models", x = "Model", y = "FDR") + 
     guides(color = guide_legend(order = 1, override.aes = list(size=5)), 
              size = guide_legend(order = 2, reverse = TRUE))

@@

``` {r, echo = FALSE, message=FALSE, warning=FALSE}

models$plot_conditions = paste0(models$conditions, " Conditions")
models$plot_sd = paste0("SD ", models$sd)

png("../supplementary/sim_new/sim1_FDR.png", width = 750, height = 500)
models %>% mutate(Model = factor(Model, levels=c("Proposed", "Proposed no adjustment", "Limma", 
                                               "Limma no adjustment", "T-Test", "T-Test no adjustment")),
                  Replicates = as.factor(rep)) %>%
  filter(Model %in% c("Proposed", "T-Test", "Limma")) %>% 
  ggplot() + geom_point(aes(x = Model, y = fpr, size = Replicates, color = Model)) + 
  scale_colour_manual(values=cbPalette) +
  scale_size_manual(values = c(3,5,7,9)) + 
  facet_grid(vars(plot_sd), vars(plot_conditions)) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 60, hjust=1, size = 16), 
         axis.text.y = element_text(size = 16), 
         legend.text=element_text(size=18),
         axis.title.y = element_text(size = 22),
         axis.title.x = element_text(size = 22),
         title = element_text(size = 22),
         strip.text = element_text(size = 16)) + 
  labs(title = "Simulation 1: FDR adjusted models", x = "Model", y = "FDR") + 
   guides(color = guide_legend(order = 1, override.aes = list(size=5)), 
              size = guide_legend(order = 2, reverse = TRUE))
dev.off()

# axis.text.x = element_text(angle = 60, hjust=1, size = 16),
#         axis.text.y = element_text(size = 16),
#         legend.text=element_text(size=18),
#         axis.title.y = element_text(size = 22),
#         axis.title.x = element_text(size = 22),
#         title = element_text(size = 22),
#         strip.text = element_text(size = 16),

@@

```{r echo=FALSE, message=FALSE, warning=FALSE}

models$plot_conditions = paste0(models$conditions, " Conditions")
models$plot_sd = paste0("SD ", models$sd)

png("../supplementary/sim_new/sim1_Recall.png", width = 750, height = 500)
models %>% mutate(Model = factor(Model, levels=c("Proposed", "Proposed no adjustment", "Limma", 
                                               "Limma no adjustment", "T-Test", "T-Test no adjustment")),
                  Replicates = as.factor(rep)) %>%
  filter(Model %in% c("Proposed", "T-Test", "Limma")) %>% 
  ggplot() + geom_point(aes(x = Model, y = recall, size = Replicates, color = Model)) + 
  scale_colour_manual(values=cbPalette) +
  scale_size_manual(values = c(3,5,7,9)) + 
  facet_grid(vars(plot_sd), vars(plot_conditions)) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 60, hjust=1, size = 16), 
         axis.text.y = element_text(size = 16), 
         legend.text=element_text(size=18),
         axis.title.y = element_text(size = 22),
         axis.title.x = element_text(size = 22),
         title = element_text(size = 22),
         strip.text = element_text(size = 16)) + 
  labs(title = "Simulation 1: Recall adjusted models", x = "Model", y = "Recall") + 
   guides(color = guide_legend(order = 1, override.aes = list(size=5)), 
              size = guide_legend(order = 2, reverse = TRUE))
dev.off()
```

@@

```{r echo=FALSE, message=FALSE, warning=FALSE}
png("../supplementary/sim_new/sim1_accuracy.png", width = 750, height = 500)
models %>% mutate(Model = factor(Model, levels=c("Proposed", "Proposed no Adjustment", "Limma", 
                                               "Limma no Adjustment", "T-Test", "T-Test no Adjustment")),
                  Replicates = as.factor(rep)) %>%
  filter(Model %in% c("Proposed", "T-Test", "Limma")) %>% 
  ggplot() + geom_point(aes(x = Model, y = accuracy, size = Replicates, color = Model)) + 
  scale_colour_manual(values=cbPalette) +
  scale_size_manual(values = c(3,5,7,9)) + 
  facet_grid(vars(plot_sd), vars(plot_conditions)) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 60, hjust=1, size = 16), 
         axis.text.y = element_text(size = 16), 
         legend.text=element_text(size=18),
         axis.title.y = element_text(size = 22),
         axis.title.x = element_text(size = 22),
         title = element_text(size = 22),
         strip.text = element_text(size = 16)) + 
  labs(title = "Simulation 1: Accuracy adjusted models", x = "Model", y = "Accuracy") + 
   guides(color = guide_legend(order = 1, override.aes = list(size=5)), 
              size = guide_legend(order = 2, reverse = TRUE))

@@

boxplot_data$plot_conditions = paste0(boxplot_data$cond, " Conditions")
boxplot_data$plot_sd = paste0("SD ", boxplot_data$sd)

png("../supplementary/sim_new/sim1_FC_boxplot.png", width = 750, height = 500)
boxplot_data %>% mutate(Model = factor(Model, levels=c("Proposed", "Limma", "T-test")),
                        reps = factor(reps, levels = c(2,3,5,10))) %>% 
  ggplot()  + geom_hline(yintercept = .75, color = "red", size = 1.1) + 
  geom_boxplot(aes(y = log2FC, x = reps, fill = Model)) + 
  facet_grid(vars(plot_sd), vars(plot_conditions)) + 
  scale_fill_manual(values=cbPalette) +
  theme_bw() +
  theme(axis.text.x = element_text(size = 16), 
         axis.text.y = element_text(size = 16), 
         legend.text=element_text(size=18),
         axis.title.y = element_text(size = 22),
         axis.title.x = element_text(size = 22),
         title = element_text(size = 22),
         strip.text = element_text(size = 16)) + 
  labs(title = "Simulation 1: Fold Change Distribution", x = "Replicates", y = "Log2FC")
dev.off()
```

@@

load(file = "../data/ptm_models_sim2.rda")#ptm_models_sim1
load(file = "../data/adjusted_models_sim2.rda")#adjusted_models_sim1 
load(file = "../data/ttest_models_sim2.rda")#ttest_sim1
load(file = "../data/adj_ttest_models_sim2.rda")#adj_ttest_sim1
load(file = "../data/limma_models_sim2.rda")#limma_results_sim1
load(file = "../data/adj_limma_models_sim2.rda")#adj_limma_sim1

@@

``` {r, echo = FALSE, message=FALSE, warning=FALSE}
models$plot_conditions = paste0(models$conditions, " Conditions")
models$plot_sd = paste0("SD ", models$sd)

png("../supplementary/sim_new/sim3_FDR.png", width = 350, height = 350)
models %>% mutate(Model = factor(Model, levels=c("Proposed", "Proposed_no_adj", "limma_no_adj", 
                                               "Limma", "ttest_no_adj", "T-Test")),
                  Replicates = as.factor(rep)) %>%
  filter(Model %in% c("Proposed", "T-Test", "Limma")) %>% 
  ggplot() + geom_point(aes(x = Model, y = fpr, size = Replicates, color = Model)) + 
  scale_colour_manual(values=cbPalette) +
  facet_grid(vars(sd), vars(conditions)) + 
  theme_bw() +
  theme(axis.text.x = element_blank(), 
                                                 axis.text.y = element_text(size = 12), 
                                                 legend.text=element_text(size=12),
                                                 axis.title.y = element_text(size = 14),
                                                 axis.title.x = element_text(size = 14),
                                                 title = element_text(size = 16),
                                                 strip.text = element_text(size = 12)) + 
  labs(title = "FDR adjusted models", x = "Model", y = "FDR")
dev.off()
```

@@

```{r echo=FALSE, message=FALSE, warning=FALSE}
models$plot_conditions = paste0(models$conditions, " Conditions")
models$plot_sd = paste0("SD ", models$sd)

png("../supplementary/sim_new/sim3_Recall.png", width = 750, height = 500)
models %>% mutate(Model = factor(Model, levels=c("Proposed", "Proposed_no_adj", "limma_no_adj", 
                                               "Limma", "ttest_no_adj", "T-Test")),
                  Replicates = as.factor(rep)) %>%
  filter(Model %in% c("Proposed", "T-Test", "Limma")) %>% 
  ggplot() + geom_point(aes(x = Model, y = recall, size = Replicates, color = Model)) + 
  scale_colour_manual(values=cbPalette) +
  scale_size_manual(values = c(3,5,7,9)) + 
  facet_grid(vars(plot_sd), vars(plot_conditions)) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 60, hjust=1, size = 16), 
         axis.text.y = element_text(size = 16), 
         legend.text=element_text(size=18),
         axis.title.y = element_text(size = 22),
         axis.title.x = element_text(size = 22),
         title = element_text(size = 22),
         strip.text = element_text(size = 16)) + 
  labs(title = "Simulation 2: Recall adjusted models", x = "Model", y = "Recall") + 
   guides(color = guide_legend(order = 1, override.aes = list(size=5)), 
              size = guide_legend(order = 2, reverse = TRUE))
dev.off()

@@

```{r echo=FALSE, message=FALSE, warning=FALSE}
png("../supplementary/sim_new/sim3_Accuracy.png", width = 750, height = 500)
models %>% mutate(Model = factor(Model, levels=c("Proposed", "Proposed_no_adj", "limma_no_adj", 
                                               "Limma", "ttest_no_adj", "T-Test")),
                  Replicates = as.factor(rep)) %>%
  filter(Model %in% c("Proposed", "T-Test", "Limma")) %>% 
  ggplot() + geom_point(aes(x = Model, y = accuracy, size = Replicates, color = Model)) + 
  scale_colour_manual(values=cbPalette) +
  scale_size_manual(values = c(3,5,7,9)) + 
  facet_grid(vars(plot_sd), vars(plot_conditions)) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 60, hjust=1, size = 16), 
         axis.text.y = element_text(size = 16), 
         legend.text=element_text(size=18),
         axis.title.y = element_text(size = 22),
         axis.title.x = element_text(size = 22),
         title = element_text(size = 22),
         strip.text = element_text(size = 16)) + 
  labs(title = "Simulation 2: Accuracy adjusted models", x = "Model", y = "Accuracy") + 
   guides(color = guide_legend(order = 1, override.aes = list(size=5)), 
              size = guide_legend(order = 2, reverse = TRUE))
dev.off()
```

@@

boxplot_data$plot_conditions = paste0(boxplot_data$cond, " Conditions")
boxplot_data$plot_sd = paste0("SD ", boxplot_data$sd)

png("../supplementary/sim_new/sim3_FC_boxplot.png", width = 750, height = 500)
boxplot_data %>% mutate(Model = factor(Model, levels=c("Proposed", "Limma", "T-test")),
                        reps = factor(reps, levels = c(2,3,5,10))) %>% 
ggplot()  + geom_hline(yintercept = .75, color = "red", size = 1.1) + 
geom_boxplot(aes(y = log2FC, x = reps, fill = Model)) + facet_grid(vars(sd), vars(cond)) + 
    scale_fill_manual(values=cbPalette) +
  theme_bw() +
  theme(axis.text.x = element_text(size = 16), 
         axis.text.y = element_text(size = 16), 
         legend.text=element_text(size=18),
         axis.title.y = element_text(size = 22),
         axis.title.x = element_text(size = 22),
         title = element_text(size = 22),
         strip.text = element_text(size = 16)) + 
labs(title = "Simulation 2: Fold Change Distribution", x = "Replicates", y = "Log2FC")
dev.off()
```
